---
title: "vis_and_eda"
output: html_document
---
```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Making a first plot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()

```

same as the last one
```{r}
weather_df |>
  ggplot(aes(x=tmin,y=tmax))+
  geom_point()
```

```{r}
gg_weather_scateerplot = 
  weather_df |>
  ggplot(aes(x=tmin,y=tmax))+
  geom_point()

ggp_weather_scatterplot
```

```{r}
weather_df |>
  filter(is.na(tmax))
```

##fancier scatterpot
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color =name))+
  geom_point(alpha = .3, size = 0.8) +
  geom_smooth(se = FALSE)
```


where you define aesthetis can matter
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax))+
  geom_point(aes(color = name),alpha = .3, size = 0.8) +
  geom_smooth(se = FALSE)

```

use faceting real quick
```{r}
weather_df |>
  ggplot(aes(x=tmin, y =tmax, color = name))+
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~.)
```


making a somewhat more interesting scatterplot
```{r}
weather_df |>
  ggplot(aes(x=date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3)+
  geom_smooth(se = FALSE)+
  facet_grid(.~ name)

```

Learning assessment
```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax_fahr = tmax *(9/5) +32,
    tmin_fahr = tmin *(9/5) + 32
  ) |>
  ggplot(aes(x=tmin_fahr, y = tmax_fahr)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```


##small things 
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax))+
  geom_smooth(se = FALSE)
```

```{r}
weather_df |>
   ggplot(aes(x = tmin, y = tmax))+
   geom_hex()
```

```{r}
weather_df |>
   ggplot(aes(x = tmin, y = tmax, color = "blue"))+
   geom_point()
```


```{r}
weather_df |>
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram(position = "dodge")

```


fixxing it by facet?
```{r}
weather_df |>
  ggplot(aes(x =tmin, fill = name))+
  geom_histogram()+
  facet_grid(. ~ name)

```

by density plot?
```{r}
weather_df |>
  ggpplot(aes(x = tmin, fill = name))+
  geom_density(alpha = .3)

```


boxplot
```{r}
weather_df |>
  ggpplot(aes(x = name, y = tmin, fill = name))+
  geom_boxplot()
```

violin plots
```{r}
weather_df |>
  ggpplot(aes(x = name, y = tmin, fill = name))+
  geom_violin()


```


ridge plot
```{r}
weather_df |>
  ggpplot(aes(x = tmin, y = name))+
  geom_density_ridges()
```


## learning assessment 2
```{r}
weather_df |>
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) 

weather_df |>
  ggplot(aes(x = name, y = prcp)) + 
  geom_boxplot()

weather_df |>
  filter(prcp >10, prcp <1000) |>
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .3)
```


## Saving plots
```{r}
weather_df |>
  ggplot(aes(x= date, y = tmax, color = name))+
  geom_point()

ggsave("ggp_weather.pdf")

ggp_weather_df |>
  ggplot(aes(x= date, y = tmax, color = name))+
  geom_point()

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 6)
```


Embedding plots
```{r,fig.width = 6, fig.asp = .6}
weather_df |>
  ggplot(aes(x= date, y = tmax, color = name))+
  geom_point()


```



make a scaterplot but fancy this time
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)
  lab(
    title= "Temperate scatterplot", 
    x = "Minimum Temp (C)",
    x = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  )+
    scale_x_continuous(
      breaks = c(-15,0,20),
      labels = c("-15","0","20")
    )+
    scale_y_continuous(
      limits = c(0,30),
      transform = "sqrt"
    )
```

look at color:
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)
  lab(
    title= "Temperate scatterplot", 
    x = "Minimum Temp (C)",
    x = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  )+
    viridis::scale_color_viridis(discrete = TRUE)
```


## Themes
```{r}
ggp_scatterplot = 
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)
  lab(
    title= "Temperate scatterplot", 
    x = "Minimum Temp (C)",
    x = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  )+
    viridis::scale_color_viridis(discrete = TRUE)

```

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom")
```


```{r}
ggp_scatterplot +
  theme_bw()+
  theme(legend.position = "bottom")
```


```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
theme_bw()
```


```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
theme_minimal()

```

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
theme_classic()
```


```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  ggthemes::theme_excel()
```

learning assessment
```{r}
weather_df |>
ggplot(aes(x = date, y = tmax, color = name))+
  geom_point()+
  geom_smooth(se=FALSE)+
  viridis::scale_color_viridis(discrete = TRUE)+
  labs(
    x = "Date",
    y = "Maximum Temp(C)",
    title = "Seasonal variation in Max Temp"
  )+
  theme_minimal()+
  theme(legend.position = "bottom")

```

extra bonus stuff in 'ggplot'
use different datasets in different 'geom's
```{r}
central_park_df =
  weather_df |>
  filter(name == "Central_NY")

molokai_df |>
  filter(name =="Molokai_HI")

molokai_df |>
  ggplot(aes(x=data, y = tmax, color = name))+
  geom_point()
geom_line(data = central_park_df)
```

## multiple panels
```{r}
weather_df |>
  ggplot(aes(x = tmax, fill = name))+
  geom_density()+
  facet_grid(.~name)
```

```{r}
ggp_tmax_tmin =
  weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name))+
  geom_point(alpha =.3)

ggp_tmax_density=
  weather_df |>
  ggplot(aes(x=tmax, fill = name))+
  geom_density(alpha=.3)

ggp_tmax_date = 
  weather_df |>
  ggplot(aes(x=date, y = tmax, color = name))+
  geom_point()+
  geom_smooth(se = FALSE)

(ggp_tmax_density + ggp_tmax_tmin)/ggp_tmax_date
```


